---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---


```{r echo=FALSE, include=FALSE}
library(CDM)
```

# Data

## Assessment Data

```{r}

str(data.ecpe$data)

```


## Q Matrix:

```{r}

str(data.ecpe$q.matrix)

```

# Model

## DINA
```{r}

ecpe.dina <- din(data.ecpe$data[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina
```

### Guess parameters
```{r}
str(ecpe.dina$guess)

```

### Slip parameters
```{r}
str(ecpe.dina$slip)

```

### Coefficients, Standard errors and Confidence intervals for all parameters 
```{r}

param <- IRT.se(ecpe.dina, extended=TRUE) 
head(param)
```


```{r}

tail(param, 15)

```

Let's separate each type of parameter

```{r}

p <- split(param, param$partype)

```

### Item characteristics

#### Item p-values

```{r}

pvalues <- colMeans(data.ecpe$data[,-1], na.rm=TRUE) 
pvalues
```

#### Guessing parameter

```{r}

head(p$guess)

```

#### Slipping parameter

```{r}

head(p$slip)

```

#### Item Discrimination parameter

```{r}

omega1 <- 1 - p$guess$est - p$slip$est 

omega1 

```

#### Item Easiness parameter

```{r}

omega2 <- (p$guess$est + (1 - p$slip$est))/2 

omega2 

```

### Skills Characteristics

#### Skills Distribution

```{r}

p$margprobs

```

#### Skills Class Distribution

```{r}

p$probs

```

#### Individual Skills profile

```{r}
skill.p <- IRT.factor.scores(ecpe.dina, type="MLE") 
head(skill.p)
```

```{r}
summary(ecpe.dina)
```

```{r}
print(ecpe.dina)
```

```{r}
par(mfrow=c(2,2)) 
plot(ecpe.dina)
```


```{r}
# parameters and standard errors of DINA model 
param <- IRT.se(ecpe.dina, extended=TRUE) 
p <- split(param, param$partype)
p
```

```{r}
# items characteristics (cf. Table 6) 
pvalues <- colMeans(data.ecpe$data[,-1], na.rm=TRUE) # item p-values 
p$guess # guessing parameters 
p$slip # slipping parameters 
omega1 <- 1 - p$guess$est - p$slip$est # item discrimination 
omega2 <- (p$guess$est + (1 - p$slip$est))/2 # item easiness

```


```{r}

# skill characteristics (cf. Table 7 and 8) 
p$margprobs # skill distribution Q1 
p$probs # skill class distribution Q2 
IRT.factor.scores(ecpe.dina, type="MLE")[1:5,] # individual skill profile Q3
```

```{r}
# plot model parameters 
par(mfrow=c(2,2)) 
plot(ecpe.dina, pattern=data.ecpe$data[1,-1])

```


```{r}

# correlation between skills 
skill.cor(ecpe.dina)$cor.skills
# various fit criteria 
fit.ecpe <- IRT.modelfit(ecpe.dina)
```

```{r}

# new Q-matrices 
newq13 <- newq23 <- data.ecpe$q.matrix 
newq13[,4] <- 1*(newq13[,1]==1 | newq13[,3]==1) 
newq23[,4] <- 1*(newq23[,2]==1 | newq23[,3]==1) 
newq13 <- newq13[,c(2,4)] 
newq23 <- newq23[,c(1,4)]
```

```{r}

# define, estimate and derive model fit of competing models 
ecpe13 <- din(data.ecpe$data[,-1], newq13) 
fit.ecpe13 <- IRT.modelfit(ecpe13) 
ecpe23 <- din(data.ecpe$data[,-1], newq23) 
fit.ecpe23 <- IRT.modelfit(ecpe23)
```

```{r}
# compare competing models 
IRT.compareModels(fit.ecpe, fit.ecpe13, fit.ecpe23)


```

