---
title: "DINA model on ECPE Dataset"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---


```{r echo=FALSE, include=FALSE}
library(CDM)
library(dplyr)
```

# Data

## Assessment Data

```{r}

str(data.ecpe$data)
```

```{r}
data.ecpe.sample_10 = sample_n(data.ecpe$data, size = 10)
data.ecpe.sample_20 = sample_n(data.ecpe$data, size = 20)
data.ecpe.sample_30 = sample_n(data.ecpe$data, size = 30)
data.ecpe.sample_33 = sample_n(data.ecpe$data, size = 30) # 10 % more than total number of variables
data.ecpe.sample_36 = sample_n(data.ecpe$data, size = 36) # 20 % more than total number of variables
data.ecpe.sample_45 = sample_n(data.ecpe$data, size = 45) # 50 % more than total number of variables
data.ecpe.sample_60 = sample_n(data.ecpe$data, size = 60) # 100 % more than total number of variables
data.ecpe.sample_100 = sample_n(data.ecpe$data, size = 100) # 100 % more than total number of variables
str(data.ecpe.sample_30)
```


## Q Matrix:

```{r}

str(data.ecpe$q.matrix)

```

# Model

## DINA
```{r}

ecpe.dina <- din(data.ecpe$data[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina
```

### Generate model for different sample size

```{r}

ecpe.dina.sample_10 <- din(data.ecpe.sample_10[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_10
```

```{r}

ecpe.dina.sample_20 <- din(data.ecpe.sample_20[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_20
```

```{r}

ecpe.dina.sample_30 <- din(data.ecpe.sample_30[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_30
```

```{r}

ecpe.dina.sample_33 <- din(data.ecpe.sample_33[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_33
```

```{r}
ecpe.dina.sample_36 <- din(data.ecpe.sample_36[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_36
```

```{r}
ecpe.dina.sample_45 <- din(data.ecpe.sample_45[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_45
```

```{r}
ecpe.dina.sample_60 <- din(data.ecpe.sample_60[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_60
```


```{r}
ecpe.dina.sample_100 <- din(data.ecpe.sample_100[,-1], data.ecpe$q.matrix,progress = FALSE)
ecpe.dina.sample_100
```

### Compare DINA models with different sample size

```{r}

ecpe.dina.full_2922 <- ecpe.dina

compare.model <- IRT.compareModels(ecpe.dina.sample_10, ecpe.dina.sample_20, ecpe.dina.sample_30, ecpe.dina.sample_33, ecpe.dina.sample_36, ecpe.dina.sample_45, ecpe.dina.sample_60, ecpe.dina.sample_100, ecpe.dina.full_2922)


compare.model$IC[,'Sample Size'] <- c(10, 20, 30, 33, 36, 45, 60, 100, 2922)
compare.model$IC[,'AIC_normalized'] <- compare.model$IC[,'AIC'] / c(10, 20, 30, 33, 36, 45, 60, 100, 2922) 
compare.model$IC[,'loglike_normalized'] <- compare.model$IC[,'loglike'] / c(10, 20, 30, 33, 36, 45, 60, 100, 2922) 

compare.model$IC[,c('Model', 'Sample Size', 'AIC_normalized')] # c('Model','loglike_normalized', 'AIC_normalized')
compare.model$IC[,c('Model', 'Sample Size', 'loglike', 'AIC')]
```

### Get model fit

```{r}

fit.ecpe <- IRT.modelfit(ecpe.dina)
fit.ecpe.sample_30 <- IRT.modelfit(ecpe.dina.sample_30)
fit.ecpe.sample_33 <- IRT.modelfit(ecpe.dina.sample_33)
fit.ecpe.sample_36 <- IRT.modelfit(ecpe.dina.sample_36)
fit.ecpe.sample_45 <- IRT.modelfit(ecpe.dina.sample_45)
fit.ecpe.sample_60 <- IRT.modelfit(ecpe.dina.sample_60)
fit.ecpe.sample_100 <- IRT.modelfit(ecpe.dina.sample_100)

```

```{r}

IRT.compareModels(fit.ecpe, fit.ecpe.sample_30, fit.ecpe.sample_33, fit.ecpe.sample_36, fit.ecpe.sample_45, fit.ecpe.sample_60, fit.ecpe.sample_100)

```

## Details of DINA model

### Guess parameters
```{r}
str(ecpe.dina$guess)

```

### Slip parameters
```{r}
str(ecpe.dina$slip)

```

### Coefficients, Standard errors and Confidence intervals for all parameters 
```{r}

param <- IRT.se(ecpe.dina, extended=TRUE) 
head(param)
```


```{r}

tail(param, 15)

```

**Let's create separate copies of each type of parameter** 

```{r}

p <- split(param, param$partype)

```

### Item characteristics

#### Item p-values

```{r}

pvalues <- colMeans(data.ecpe$data[,-1], na.rm=TRUE) 
pvalues
```

#### Guessing parameter

```{r}

head(p$guess)

```

#### Slipping parameter

```{r}

head(p$slip)

```

#### Item Discrimination parameter

```{r}

omega1 <- 1 - p$guess$est - p$slip$est 

omega1 

```

#### Item Easiness parameter

```{r}

omega2 <- (p$guess$est + (1 - p$slip$est))/2 

omega2 

```

### Skills Characteristics

#### Skills Distribution

```{r}

p$margprobs

```

#### Skills Class Distribution

```{r}

p$probs

```

#### Individual Skills profile

```{r}
skill.p <- IRT.factor.scores(ecpe.dina, type="MLE") 
head(skill.p)
```

## Summarizing DINA model

```{r}
summary(ecpe.dina)
```

```{r}
print(ecpe.dina)
```

```{r}
par(mfrow=c(2,2)) 
plot(ecpe.dina)
```

